Ecosystem size and complexity are extrinsic drivers of food chain length in branching networks

### 2022-03-31 rmd/manuscript.Rmd
edit: manuscript

```{=tex}
\begin{align}
W_{B,x}(t) = W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
\end{align}
```
where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of intraguild predator to basal species over intraguild prey. We assume the Holling Type-II functional response for predator-prey interactions:

@@

The population growth of the intraguild prey is controlled by conversion efficiency, number of prey captured, background mortality, and predation:

@@

The population of the intraguild predator is fueled by two prey species:

@@

***Food chain length.*** At each patch and time step, we determined FCL  as the highest trophic position ($TP_{x}(t)$) of the existing species. When the intraguild predator exists, the trophic position is determined as follows:

@@

Finally, we investigated three levels of omnivory (**Figure 1D**). Trophic interaction strengths can be controlled by varying the attack rate and/or conversion efficiency [@wardMechanisticTheoryAquatic2017]. We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for intraguild predator $P$. Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$. To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.

@@

>>>>

```{=tex}
\begin{align}
W_{B,x}(t) = W_{BC,x}(t) + \delta_{x}(t) W_{BP,x}(t)
\end{align}
```
where $W_{ij,x}(t)$ is the population-wide consumption of consumer $j$ on prey $i$ (number of individuals), and the parameter $\delta_{x}(t)$ is the relative preference of predator to basal species over consumer. We assume the Holling Type-II functional response for predator-prey interactions:

@@

The population growth of the consumer is controlled by conversion efficiency, number of prey captured, background mortality, and predation:

@@

The population of the predator is fueled by two prey species:

@@

***Food chain length.*** At each patch and time step, we determined FCL as the highest trophic position ($TP_{x}(t)$) of the existing species. When the predator exists, the trophic position is determined as follows:

@@

Finally, we investigated three levels of omnivory (**Figure 1D**). Trophic interaction strengths can be controlled by varying the attack rate and/or conversion efficiency [@wardMechanisticTheoryAquatic2017]. We changed attack rate ($a_{BP}$ and $a_{CP}$) and conversion efficiency ($e_{BP}$ and $e_{CP}$) of two prey items for predator $P$. Specifically, we used the following combinations to represent three levels of omnivory: "Chain", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[1], max(sim_main_result$e_cp))` $\}$; "Weak omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(min(sim_main_result$a_bp), max(sim_main_result$a_cp), unique(sim_main_result$e_bp)[2], max(sim_main_result$e_cp))` $\}$; and "Strong omnivory", $\{a_{BP}, a_{CP}, e_{BP}, e_{CP}\}=\{$ `r c(max(sim_main_result$a_bp), min(sim_main_result$a_cp), unique(sim_main_result$e_bp)[3], min(sim_main_result$e_cp))` $\}$. To simplify simulation setup, we fixed the following parameters: $e_{BC}=$ `r unique(sim_main_result$e_bc)` , $a_{BC}=$ `r unique(sim_main_result$a_bc)`, and $h_{ij}=$ `r unique(sim_main_result$h_bc)`.

@@

# **Acknowledgements**

We thank C Perovich for thoughtful discussions in the early stages of this project. This material is based upon work supported by the National Science Foundation through the Division of Environmental Biology (DEB 2015634).

# **Author contributions**

JP developed the theoretical model of trophic dynamics, ran the simulations, and made the figures with input from AT. JP wrote the first draft, while AT and JF contributed significantly to the final version.

# **Data availability**

No data were used or produced in this research. R functions for simulations are provided as the R package *mcbrnet* (available at <https://github.com/aterui/mcbrnet>). Codes for simulations, statistical analysis, figures, and tables are available at <https://github.com/aterui/public-proj_fcl-river-theory>.
